---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: true
  pdf_document: default
---
```{r setoptions, echo=TRUE}
knitr::opts_chunk$set(echo=TRUE,results="show")
```

# Peer Assessment 1

## **Loading and preprocessing the data**
```{r load_data}
library(readr)
activity <- read_csv("activity.csv")
```

## Here we calculate the total number of steps per day
```{r total_step}
library(dplyr)
activity_steps<-activity %>%
  group_by(date)%>%
  summarise(total_steps=sum(steps, na.rm=TRUE))
activity_steps
```

## What is mean total number of steps taken per day? The mean number of steps per day is 9354, while the median is 10395.
```{r summary1}
summary(activity_steps$total_steps)
```

## We construct the histogram to show the distribution of the total steps per day.
```{r histogram1}
hist(activity_steps$total_steps,
     main="Total steps per day",
     xlab = "Number of steps",
     breaks = 10,
     col="magenta")
```
## The total number of steps taken per day is 570608.
```{r sum}
sum(activity_steps$total_steps)
```

## **What is the average daily activity pattern?**
## Now we create a time series plot to visualise the average steps made on 5-minute intervals.
```{r x}
x<- activity%>%
  group_by(interval)%>%
  summarise(averageSteps=mean(steps, na.rm=TRUE))
```

```{r plot}
with(x,plot(interval,averageSteps,
            pch=23,
            col="blue",
            type="l",
            xlab="5-minute intervals",
            ylab="Average steps",
            main="Average daily activity pattern"))
```

## the 5-minute interval with the most number of steps is 835, witg an average of 206 steps.

```{r max_intervap}
max_interval<- x[which.max(x$averageSteps), ]
max_interval
```

## **Imputing missing values**
## We have a total of 2304 NA values in the activity dataset.
```{r NA}
sum(is.na(activity))
```

## So in order to fill those NAs, we merge the data set grouped by intervals with the main dataset.
```{r data_merge}
data_filled<-merge(activity,x,by="interval")
data_filled$steps[is.na(data_filled$steps)]<-data_filled$averageSteps
```
## And we check that the new dataset has indeed 0 NAs.
```{r sum2}
sum(is.na(data_filled))
```

## Now we are creating a histogram of the total number of steps taken each day, without any NAs.We notice that the plot looks identical as the plot with the NA values.
```{r data_filled}
data_filled2<-data_filled%>%
group_by(date)%>%
summarize(total_steps2=sum(steps,na.rm=TRUE))
```
```{r histogram2}
hist(data_filled2$total_steps2,
     main="Total steps per day (without NAs)",
     xlab = "Number of steps",
     breaks = 10,
     col="magenta")
```

## The median is also the same (10395), but the mean is a little higher, namely 9371. This indicates that the NA values skew the results and make them less accurate. In this case, though, the difference isn't so significant.  
```{r summary2}
summary(data_filled2$total_steps2)
```

## **Are there differences in activity patterns between weekdays and weekends?**
```{r weekends_weekdays}
library(dplyr)

unique_days<-unique(data_filled$date)
day_names<-weekdays(as.Date(unique_days))
day_names

data_filled<-data_filled%>%
  mutate(day_name = weekdays(as.Date(date)),
         day_type = ifelse(day_name %in% c("Saturday", "Sunday"), "weekend", "weekday"))  

data_weekdays<-data_filled%>%
  group_by(interval,day_type)%>%
  summarise(avgSteps=mean(steps))
```

```{r plot2}
library(ggplot2)
ggplot(data = data_weekdays, aes(interval,avgSteps))+
  geom_line()+
  facet_grid(day_type~.)+
  labs(x = "5-minute interval",
       y="Average number of steps")
```

## we can observe that, whereas during weekdays the average number of steps peaks in the 570-1000 minute intervals and then decreases for the rest of the intervals, during the weekends the avg number of steps seems more consistent across intervals, without outstanding peaks.

git remote set-url origin git@github.com:caty-g15/RepData_PeerAssessment1.git


